cardinalR: Generating interesting high-dimensional data structures

### 2025-09-16 paper-cardinalR.Rmd
Updated gen_linearbranches text by adding slope and intercept info

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches $1$ and $2$ are initialized with fixed slopes and intercepts. Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. Branches $3$ to $k$ are added iteratively. Each additional branch $i$ starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. The $X_1$ values are defined over a short range, from $x_{start}$ to $x_{start} + 1$. The $X_2$ value is calculated using the formula $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$, where $s_i$ is a chosen slope from a selected branch, and $\epsilon \sim U(0, 0.2)$. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.

>>>>

The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points in $p-D$ forming $k$ linear branches. Each branch is a line segment with added jitter to simulate noise. Branches $1$ and $2$ are initialized with fixed slopes and intercepts: branch $1$ has slope $0.5$ and intercept $0$, with $X_1 \sim U(-2, 8)$ and $X_2 = 0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$; branch $2$ has slope $-0.5$ and intercept $0$, with $X_1 \sim U(-6, 2)$ and $X_2 = -0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$.

Branches $3$ to $k$ are added iteratively. Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches. Specifically, $X_1$ starting points are sampled outside the intervals $[-8, -7]$, $[-2, 2]$, and $[7, 8]$, and $X_2$ starting points are sampled outside $[7, 8]$. For each new branch $i$, $X_1$ values are drawn over a short interval $[x_{\text{start}}, x_{\text{start}} + 1]$, and $X_2$ values are calculated as $X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon$, where $s_i$ is a randomly chosen slope from a filtered sequence and $\epsilon \sim U(0, 0.2)$.

For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, for $j = 3, \dots, p$.

### 2025-09-16 paper-cardinalR.tex
Updated gen_linearbranches text by adding slope and intercept info

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts. Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). Branches \(3\) to \(k\) are added iteratively. Each additional branch \(i\) starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. The \(X_1\) values are defined over a short range, from \(x_{start}\) to \(x_{start} + 1\). The \(X_2\) value is calculated using the formula \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\), where \(s_i\) is a chosen slope from a selected branch, and \(\epsilon \sim U(0, 0.2)\). For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).

>>>>

The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points in \(p-D\) forming \(k\) linear branches. Each branch is a line segment with added jitter to simulate noise. Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts: branch \(1\) has slope \(0.5\) and intercept \(0\), with \(X_1 \sim U(-2, 8)\) and \(X_2 = 0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\); branch \(2\) has slope \(-0.5\) and intercept \(0\), with \(X_1 \sim U(-6, 2)\) and \(X_2 = -0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\).

Branches \(3\) to \(k\) are added iteratively. Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches. Specifically, \(X_1\) starting points are sampled outside the intervals \([-8, -7]\), \([-2, 2]\), and \([7, 8]\), and \(X_2\) starting points are sampled outside \([7, 8]\). For each new branch \(i\), \(X_1\) values are drawn over a short interval \([x_{\text{start}}, x_{\text{start}} + 1]\), and \(X_2\) values are calculated as \(X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon\), where \(s_i\) is a randomly chosen slope from a filtered sequence and \(\epsilon \sim U(0, 0.2)\).

For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), for \(j = 3, \dots, p\).
